---
title: "STA457 assignment"
author: "Haoluan Chen"
date: "3/12/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(readxl)
library(MASS)
library(forecast)
library(timeSeries)
library(timeDate)
library(sarima)
```
```{r functions, echo= FALSE}
# =============================================#======
# utility functions --prewhitening and LjungBox test
# Author: Jen-Wen Lin
#=====================================================

# PreWhiten.ar   
# Prewhitening time series x based on a fitted AR model
# x: Time series to be transformed (n x 1 vector)
# ar: Autoregressive coefficient estimates (p-1 x 1 vector)
PreWhiten.ar<- function(x , ar = NULL){
        if(is.null(ar)) print(" autoregressive coefficients are empty!")
        pwData = numeric(0)
        pwData = filter(x, c(1, -ar),method=c("convo"),sides=1) 
        pwData[!is.na(pwData)]
}

# PreWhiten.arma   
# Prewhitening time series x based on a fitted ARMA model
# x: Time series to be transformed (n x 1 vector)
# ar: Autoregressive coefficient estimates (p x 1 vector)
# ma: Moving-average coefficient estimates (q x 1 vector)
PreWhiten.arma<- function(x , ar = NULL, ma = 0){
        if(is.null(ar) && is.null(ma)) print("both ar and ma coefficients are empty!")
        pwData = numeric(0)
        m = as(modelCoef(new("ArmaModel", ar = ar, ma = ma)), "list")
        eps = numeric(length(x))
        pwData = xarmaFilter(m, x =x, eps = eps, whiten = TRUE) 
        pwData[!is.na(pwData)]
}

# LBTest   
# Return thet p-values of Ljung-Box portmanteau tests for a fitted ARMA model
# res: residuals from a fitted ARMA model  (n x 1 vector)
# nPQ: # of model parameters (integer)
# m: test inputs (integer)
LBTest<- function(res, nPQ = 0, m = 24, ifPlot = FALSE){
        stopifnot(nPQ >= 0, m >= 1, m > nPQ)
        n <- length(res)
        lags <- 1:m
        df <- (nPQ+1):m 
        ra <- (acf(res, lag.max = m, plot = FALSE)$acf)[-1]
        QQ <- n * (n + 2) * cumsum((ra^2)/(n - (1:m)))[df]
        
        pv <- 1 - pchisq(QQ, df)
        QQ <- round(QQ, 2)
        a <- matrix(c(df, QQ, pv), ncol = 3)
        dimnames(a) <- list(rep("", length(QQ)), c("m", "Qm", "pvalue"))
        if(ifPlot){
                plot(x = a[,1],y = a[,3],
                     ylim = c(0,1), pch = 15, col =4,
                     ylab = "p-value", xlab = "m",
                     main = "Ljung-Box portmanteau test")
                abline(h =0, col =2)
                grid()
        }else {
                a
        }
}

```

  
```{r data}
GFluTrends<-read_excel("case_study_1_fluwatch.xlsx",sheet="Google Flu Trends", skip = 1)
fluWatch<-read_excel("case_study_1_fluwatch.xlsx", sheet="FluWatch-BC", skip = 2)  

tim<-timeSequence(from = "2003-09-28", to = "2015-08-09", by = "week") 

tim1<-timeSequence(from = "2003-09-07", to = "2015-08-23", by = "week") 
GFT<- timeSeries(GFluTrends[,"British Columbia"], charvec = tim)     
fluTest<- timeSeries(fluWatch[,"FluTest"], charvec = tim1) 

# Training and test sample
GFT.obs = window(GFT, start = "2003-09-28", end = "2014-08-10")
GFT.test = window(GFT, start = "2014-08-17", end = "2015-08-09")


fluTest.obs = window(fluTest, start = "2003-09-28", end = "2014-08-10")
fluTest.test = window(fluTest, start = "2014-08-17", end = "2015-08-09")
```


```{r model selection and adequacy}
mod.arma<-auto.arima(GFT.obs, max.p = 52, max.q = 52, stationary = TRUE) 
p = mod.arma$arma[1]; q = mod.arma$arma[2]

coef(mod.arma)
plot(mod.arma)
npq = sum(mod.arma$arma[c(1,2)])
LBTest(mod.arma$residuals, nPQ = npq, m = 52, ifPlot = TRUE)
```

As shown above, the p-values of the test statistics are well above the 5% significance level so the fitted model is adequacy.

```{r Google Fle Trend lag selection}
mod = mod.arma;nAR = mod$arma[1]; nMA = mod$arma[2]

if(nMA!=0){
  xf = PreWhiten.arma(GFT.obs, ar = mod$coef[1:nAR], 
                      ma = mod$coef[(1:nMA)+nAR])[-(1:nAR)]
  yf = PreWhiten.arma(fluTest.obs, ar = mod$coef[1:nAR], 
                      ma=mod$coef[(1:nMA)+nAR])[-(1:nAR)]  
}else{
  xf = PreWhiten.arma(GFT.obs, ar = mod$coef[1:nAR], 
                      ma = 0)[-(1:nAR)]
  yf = PreWhiten.arma(fluTest.obs, ar = mod$coef[1:nAR], 
                      ma=0)[-(1:nAR)] 
}

par(cex=0.75,bg="gray95")
ccf(c(xf), c(yf), lwd=1, ylab="Cross-correlation functions",
    main="CCF of prewhitened GFT and flu test")
abline(v=0, col="gold", lwd=2, lty="dashed")
text(-1, 0.2, "-1", col=2)
```

As shown above, we will include $GFT_t$ and $GFT_{t_1}$ in our transfer function noise model.


```{r TFN model }
y<-fluTest
x<-GFT
dat<- cbind(y,x, lag(x))[-c(1:4),]

tim2<-timeSequence(from = "2003-10-05", to = "2015-08-23", by = "week") 
colnames(dat)<-c("fluTest", "GFT", "GFT1")
data<- timeSeries(dat, charvec = tim2)

data.obs = window(data, start = "2003-10-05", end = "2014-08-10")
data.test = window(data, start = "2014-08-17", end = "2015-08-09")


mod.tfn = auto.arima(data.obs[,1], xreg = data.obs[,-1], stationary = TRUE)
mod.tfn
coef(mod.tfn)

```
Here is the fitted TFN model
$$FluTest = -94.1375 + 0.0777GFT_t + 0.0587GFT_{t-1} + \eta_t$$
$$\eta_t \sim ARMA(1,1)$$
$$\eta_t - 0.829 \eta_{t-1} = a_t + 0.3667 a_{t-1}$$

```{r adequacy}
m = 26
lags = 1:m
df <- (3):m
n = length(mod.tfn$res)
rccf = ccf(mod$residuals,mod.tfn$residuals, plot = FALSE, lag.max = m)$acf[-(1:m)]
Qm = n* (n + 2) * cumsum((rccf^2)/(n - (0:m)))[df]
pv <- 1 - pchisq(Qm, df)
a = cbind(df, Qm,pv)

par(mfrow = c(1,2))
LBTest(mod.tfn$res, nPQ = 6, ifPlot = TRUE)
plot(x = a[,1],y = a[,3],
     ylim = c(0,1), pch = 15, col =4,
     ylab = "p-value", xlab = "m",
     main = "Cross-correlation check")
abline(h =0.05, col =2)
grid()
```

From above plot, the p-values of the test statistics are well above the 5% significance level so the fitted model(TFN) is adequacy.

```{r TFN model plot}
par(mfrow = c(1,1), cex = 0.75)
ts.plot(mod.tfn$fitted, ylab = "", main ="TFN model")
lines(c(fluTest.obs), pch = 10, col = "green", type ="p")
grid()


```

```{r arima on flutest}
flu.arima<- auto.arima(fluTest.obs, max.p = 52, max.q = 52,stationary = TRUE)

coef(flu.arima)
par(mfrow = c(1,2))
LBTest(flu.arima$res, nPQ = 6, ifPlot = TRUE)
plot(x = a[,1],y = a[,3],
     ylim = c(0,1), pch = 15, col =4,
     ylab = "p-value", xlab = "m",
     main = "Cross-correlation check")
abline(h =0.05, col =2)
grid()

```

From above plot, the p-values of the test statistics are well above the 5% significance level so the fitted model(arima) is adequacy.

```{r arima plot}
# arima plot
par(mfrow = c(1,1), cex = 0.75)
ts.plot(flu.arima$fitted, ylab = "", main ="Arima Plot")
lines(c(fluTest.obs), pch = 10, col = "green", type ="p")
grid()
```

```{r NN}
mod.nn = forecast::nnetar(fluTest.obs);mod.nn
mod.nnx = forecast::nnetar(data.obs[,1], xreg = data.obs[,-1]);mod.nnx
par(mfrow = c(1,1), cex = 0.75)
ts.plot(mod.nn$fitted, ylab = "", main ="Neural network")
lines(c(fluTest.obs), pch = 10, col = "green", type ="p")
grid()

par(mfrow = c(1,1), cex = 0.75)
ts.plot(mod.nnx$fitted, ylab = "", main ="NNX")
lines(c(fluTest.obs), pch = 10, col = "green", type ="p")
grid()

```
```{r forecasts for four models}

plot(forecast(flu.arima,h = 50, level = c(95)))
lines(c(rep(NA, 567), fluTest.test[-c(51,52)]),col="green", pch = 18)
grid()

plot(forecast(mod.tfn, xreg=data.test[,-1], level = (95)))
lines(c(rep(NA, 567), fluTest.test[-c(51,52)]),col="green", pch = 18)
grid()

plot(forecast(mod.nn, h = 50, level = (95)))
lines(c(rep(NA, 567), fluTest.test[-c(51,52)]),col="green", pch = 18)
grid()

plot(forecast(mod.nnx, h = 50, xreg=data.test[,-1]))
lines(c(rep(NA, 567), fluTest.test[-c(51,52)]),col="green", pch = 18)
grid()

```



```{r forecast performance in training}
accuracy(flu.arima)
accuracy(mod.tfn)
accuracy(mod.nn)
accuracy(mod.nnx)
```

In training data \newline
In ME, the TFN performs the best. \newline
In RMSE, the NNX performs the best.\newline
In MAE, the NNX performs the best.\newline
In MASE, the NNX performs the best.\newline
In ACF1 the the NN performs the best.\newline


Below are test set performance\newline

```{r h=1}


arima_forecast <- flu.arima %>%
  forecast(h = 1)
accuracy(arima_forecast, x = data.test[,1][1:1])

tfn_forecast <- mod.tfn%>%
  forecast(h =1, xreg=data.test[,-1] )
accuracy(tfn_forecast, x = data.test[,1][1:1])

nn_forecast <- mod.nn %>%
  forecast(h = 1)
accuracy(nn_forecast, x = data.test[,1][1:1])

nnx_forecast <- mod.nnx%>%
  forecast(h =1, xreg=data.test[,-1] )
accuracy(nnx_forecast, x = data.test[,1][1:1])

```
For h = 1,\newline
In ME, the ARIMA performs the best. \newline
In RMSE, the ARIMA performs the best.\newline
In MAE, the ARIMA performs the best.\newline
In MPE, the ARIMA performs the best.\newline
In MAPE, the ARIMA performs the best.\newline
In MASE, the ARIMA performs the best.\newline


```{r h=4}
arima_forecast <- flu.arima %>%
  forecast(h = 4)
accuracy(arima_forecast, x = data.test[,1][1:4])

tfn_forecast <- mod.tfn%>%
  forecast(h =4, xreg=data.test[,-1] )
accuracy(tfn_forecast, x = data.test[,1][1:4])

nn_forecast <- mod.nn %>%
  forecast(h = 4)
accuracy(nn_forecast, x = data.test[,1][1:4])

nnx_forecast <- mod.nnx%>%
  forecast(h =4, xreg=data.test[,-1] )
accuracy(nnx_forecast, x = data.test[,1][1:4])
```
For h = 4,\newline
In ME, the NN performs the best. \newline
In RMSE, the NN performs the best.\newline
In MAE, the NNX performs the best.\newline
In MPE, the NNX performs the best.\newline
In MAPE, the TFN performs the best.\newline
In MASE, the NNX performs the best.\newline


```{r h=8}
arima_forecast <- flu.arima %>%
  forecast(h = 8)
accuracy(arima_forecast, x = data.test[,1][1:8])

tfn_forecast <- mod.tfn%>%
  forecast(h =8, xreg=data.test[,-1] )
accuracy(tfn_forecast, x = data.test[,1][1:8])

nn_forecast <- mod.nn %>%
  forecast(h = 8)
accuracy(nn_forecast, x = data.test[,1][1:8])

nnx_forecast <- mod.nnx%>%
  forecast(h =8, xreg=data.test[,-1] )
accuracy(nnx_forecast, x = data.test[,1][1:8])

```
For h = 8,\newline
In ME, the ARIMA performs the best. \newline
In RMSE, the TFN performs the best.\newline
In MAE, the TFN performs the best.\newline
In MPE, the NN performs the best.\newline
In MAPE, the TFN performs the best.\newline
In MASE, the TFN performs the best.\newline


```{r h=50}
arima_forecast <- flu.arima %>%
  forecast(h = 50)
accuracy(arima_forecast, x = data.test[,1][1:50])

tfn_forecast <- mod.tfn%>%
  forecast(h =50, xreg=data.test[,-1] )
accuracy(tfn_forecast, x = data.test[,1][1:50])

nn_forecast <- mod.nn %>%
  forecast(h = 50)
accuracy(nn_forecast, x = data.test[,1][1:50])

nnx_forecast <- mod.nnx%>%
  forecast(h =50, xreg=data.test[,-1] )
accuracy(nnx_forecast, x = data.test[,1][1:50])

```
For h = 50,\newline
In ME, the TFN performs the best.\newline 
In RMSE, the TFN performs the best.\newline
In MAE, the TFN performs the best.\newline
In MPE, the ARIMA performs the best.\newline
In MAPE, the TFN performs the best.\newline
In MASE, the TFN performs the best.\newline